### 2022-09-28 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@5c6d9d4823f8bd4aab92c025a087e9b2f96261ef 🚀

Indeed, food webs are inherently low-dimensional objects, and can be
adequately represented with less than ten dimensions (Braga et al.,
2019; Eklöf et al., 2013). Simulation results by Botella et al. (2022)
suggest that there is no best method to identify architectural
similarities between networks, and that multiple approaches need to be
tested and compared to the network descriptor of interest. This matches
previous, more general results on graph embedding, which suggest that
different embedding algorithms yield different network embeddings (Goyal
\& Ferrara, 2018), calling for a careful selection of the
problem-specific approach to use. In tbl.~\ref{tbl:methods}, we present
a selection of common graph and node embedding methods, alongside
examples of their use to predict species interactions; most of these
methods rely either on linear algebra, or on pseudo-random walks on
graphs.

@@

@@

In this section, we illustrate the embedding of a collection of
bipartite networks collected by Hadfield et al. (2014), using t-SVD and
RDPG (see Strydom et al., 2022 for the full details). Briefly, an RDPG
decomposes a network into two subspaces (left and right), which are
matrices that when multiplied give an approximation of the original
network. The code to reproduce this example is available as
supplementary material (note, for the sake of comparison, that Strydom,
Catchen, et al., 2021 have an example using embedding through PCA
followed by prediction using a deep neural network on the same dataset).
The resulting (binary) metaweb \(\mathcal{M}\) has 2131 interactions
between 206 parasites and 121 hosts, and its adjacency matrix has full
rank (\emph{i.e.} it represents a space with 121 dimensions). All
analyses were done using Julia (Bezanson et al., 2017) version 1.7.2,
\emph{Makie.jl} (Danisch \& Krumbiegel, 2021), and
\emph{EcologicalNetworks.jl} (Poisot et al., 2019).

@@

@@

@@

>>>>

Indeed, food webs are inherently low-dimensional objects, and can be
adequately represented with less than ten dimensions (Braga et al.,
2019; Eklöf et al., 2013). Simulation results by Botella et al. (2022)
suggest that there is no best method to identify architectural
similarities between networks, and that multiple approaches need to be
tested and compared to the network descriptor of interest. This matches
previous, more general results on graph embedding, which suggest that
different embedding algorithms yield different network embeddings (Goyal
\& Ferrara, 2018), calling for a careful selection of the
problem-specific approach to use. In tbl.~\ref{tbl:methods}, we present
a selection of common graph and node embedding methods, alongside
examples of their use to predict species interactions; most of these
methods rely either on linear algebra, or on pseudo-random walks on
graphs. All forms of embeddings presented in the table share the common
property of summarizing their objects into (sets of) dense feature
vectors, that capture the overall network structure, pairwise
information on nodes, and emergent aspects of the network, in a
compressed way (\emph{i.e.} with some information loss, as we later
discuss in the illustration). Node embeddings tend to focus on
maintaining pairwise relationships, while graph embeddings focus on
maintaining the network structure. Nevertheless, some graph embedding
techniques (like RDPG, see \emph{e.g.} Wu et al., 2021) will provide
usable node-level embeddings.

@@

Before moving further, it is important to clarify the epistemic status
of node values derived from embeddings: specifically, they are
\emph{not} functional traits, and therefore should not be discussed in
terms of effects or responses. As per the framework of Malaterre et al.
(2019), these values neither derive from, nor result in, changes in
organismal fitness, and should therefore not be used to quantify
\emph{e.g.} functional diversity. This holds true even when there are
correlations between latent values and functional traits: although these
enable an ecological discussion of how traits condition the structure of
the network, the existence of a statistical relationship does not
elevate the latent values to the status of functional traits.

@@

In this section, we illustrate the embedding of a collection of
bipartite networks collected by Hadfield et al. (2014), using t-SVD and
RDPG (see Strydom et al., 2022 for the full details). Briefly, an RDPG
decomposes a network into two subspaces (left and right), which are
matrices that when multiplied give an approximation of the original
network. RDPG has the particularly desirable properties of being a graph
embedding technique that produces relevant node-level feature vectors,
and provides good approximations of graphs with varied structures
(Athreya et al., 2017). The code to reproduce this example is available
as supplementary material (note, for the sake of comparison, that
Strydom, Catchen, et al., 2021 have an example using embedding through
PCA followed by prediction using a deep neural network on the same
dataset). The resulting (binary) metaweb \(\mathcal{M}\) has 2131
interactions between 206 parasites and 121 hosts, and its adjacency
matrix has full rank (\emph{i.e.} it represents a space with 121
dimensions). All analyses were done using Julia (Bezanson et al., 2017)
version 1.7.2, \emph{Makie.jl} (Danisch \& Krumbiegel, 2021), and
\emph{EcologicalNetworks.jl} (Poisot et al., 2019).

@@

\leavevmode\hypertarget{ref-Athreya2017StaInf}{}%
Athreya, A., Fishkind, D. E., Levin, K., Lyzinski, V., Park, Y., Qin,
Y., Sussman, D. L., Tang, M., Vogelstein, J. T., \& Priebe, C. E.
(2017). \emph{Statistical inference on random dot product graphs: A
survey} (No. arXiv:1709.05454). arXiv.
\url{http://arxiv.org/abs/1709.05454}

@@

\leavevmode\hypertarget{ref-Malaterre2019FunDiv}{}%
Malaterre, C., Dussault, A. C., Mermans, E., Barker, G., Beisner, B. E.,
Bouchard, F., Desjardins, E., Handa, I. T., Kembel, S. W., Lajoie, G.,
Maris, V., Munson, A. D., Odenbaugh, J., Poisot, T., Shapiro, B. J., \&
Suttle, C. A. (2019). Functional Diversity: An Epistemic Roadmap.
\emph{BioScience}, \emph{69}(10), 800--811.
\url{https://doi.org/10.1093/biosci/biz089}

@@

\leavevmode\hypertarget{ref-Wu2021MaxPos}{}%
Wu, D., Palmer, D. R., \& Deford, D. R. (2021). \emph{Maximum a
Posteriori Inference of Random Dot Product Graphs via Conic Programming}
(No. arXiv:2101.02180). arXiv. \url{http://arxiv.org/abs/2101.02180}
